---
title: "moneyball"
author: "Jarrod Pelkofer"
date: "August 21, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#load libraries and data
```{r}
library(tidyverse)
library(janitor)
library(lpSolve)


batting <- read_csv("Batting.csv")
salaries <- read_csv("Salaries.csv")

batting <- batting %>%
  clean_names()

salaries <- salaries %>%
  clean_names()

```

###Add columns
```{r}
batting <- batting %>%
  mutate(ba = (h/ab),
         obp = ((h + bb + hbp)/(ab+bb+hbp+sf)),
         slg = (((h - x2b - x3b - hr) + (2* x2b) + (3*x3b) + (4*hr))/(ab)))

```

###filter batting data to 1985 and on
```{r}
batting <- batting %>%
  filter(year_id > 1984)

summary(batting)

```


###Merge Salary and Batting Data
```{r}
batting <- batting %>%
  left_join(select(salaries, everything()), by = c("player_id", "year_id"))

```

###Grab lost players from OAK A's 2001 team
```{r}

lost_players <- batting %>%
  filter(player_id %in% c("giambja01", "damonjo01", "saenzol01"), year_id == 2001) %>%
  select(player_id, h, x2b, x3b, hr, obp, slg, ba, ab)

```

###Replacement Players

Find Replacement Players for the key three players we lost! However, you have three constraints:

- The total combined salary of the three players can not exceed 15 million dollars.
- Their combined number of At Bats (AB) needs to be equal to or greater than the lost players.
- Their mean OBP had to equal to or greater than the mean OBP of the lost players

Only use the 2001 season data

```{r}
total_ab <- sum(lost_players$ab)
mean_obp <- sum(lost_players$obp)
total_sal <- 15000000

season01 <- batting %>%
  filter(year_id == 2001, !player_id %in% c("damonjo01", "giambja01", "saenzol01")) %>% #remove lost players and filter season
  drop_na() %>%
  select(year_id, player_id, salary, ab, obp)

```



###Let's solve using linear programming
```{r}

#prepare matrix for problem
mat <- matrix(0, nrow = 3, ncol = nrow(season01))
mat[1, ] <- 1 #any player can be one of the three
mat[2, ] <- season01$ab # number of ab's
mat[3, ] <- season01$salary # player salary

#Input for lp solver
objective.in <- season01$obp
const.mat <- mat
const.dir <- c("==", ">=", "<=")
const.rhs <- c(3, total_ab, total_sal)

sol <- lp(direction = "max", objective.in,  #maximize objective function
          const.mat, const.dir, const.rhs, #contraints
          all.bin = TRUE) #binary variables only

#View solution
players <- which(sol$solution == 1)
sum(season01$salary[players])
sum(season01$ab[players])
sum(season01$obp[players])

season01[players,] #optimal solution
```

